\documentclass[11pt,openany]{article}

\input{custom_commands.tex}

\setlength\parindent{0pt}

\title{The Transverse Ising Model }
\author{Mathieu Ferey}

\addbibresource{biblio.bib}


\begin{document}
	
\input{Front Page.tex}

\tableofcontents

\section{Introduction}

\section{Theoretical background}

\subsection{The $\mathds{Z}_2$ gauge theory and its Quantum Ising dual}

The Ising gauge theory, theory with a discrete $\mathds{Z}_2$ gauge symmetry in $2+1$-dimensions \cite{fradkin}

\begin{equation}
	H_{\mathds{Z}_2} = -g\sum_{\vec{x},j}\sigma^x_j(\vec{x}) - \frac{1}{g}\sum_{\vec{x}}\sigma_1^z(\vec{x})\sigma_2^z(\vec{x}+e_1)\sigma_1^z(\vec{x}+e_2)\sigma_2^z(\vec{x}),
\end{equation}

where $\vec{x}$ refers to a position on the lattice, $j=1,2$ the two possible directions of a link, $\sigma_j^{x/z}(\vec{x})$ are the Pauli matrices, living on the links of the lattice. The local operator

\begin{equation}
	Q(\vec{x}) \equiv \sigma_1^x(\vec{x})\sigma_1^x(\vec{x}-e_1)\sigma_2^x(\vec{x})\sigma_2^x(\vec{x}-e_2)
\end{equation}

commutes with $H_{\mathds{Z}_2}$. It generates local gauge transformations. One can check that $Q^2=\mathds{1}$, so that the local symmetry of our problem is indeed $\mathds{Z}_2$. The operator

\begin{equation}
	\tau^z(\vec{r}) = \prod_{(\vec{x},j) \text{ pierced by } \gamma(\vec{r})}\sigma_j^x(\vec{x}),
\end{equation}

called the magnetic charge, is a gauge invariant quantity. $\gamma$ is an open path on the dual lattice. Since

\begin{equation}
	\left\{W_p^2(\vec{r}),\tau_z^2(\vec{r})\right\} = 0\text{ and }W_p^2(\vec{r}) = \tau_z^2(\vec{r}) = \mathds{1},
\end{equation} 

one can identify $W_p$ with the Pauli matrice $\tau^x$ on  the dual lattice.

\begin{equation}
	H = -\sum_{i=1}^N\tau^z_i\tau^z_{i+1} - g\sum\tau^x_j.
\end{equation}

where $i,j$ run over the dual lattice sites.


\subsection{The classical mapping of the Quantum Ising Model}

Our transverse Ising Model is not diagonal in the eigen-basis of $S^z$ and requires a quantum treatment. Luckily for us, a combination of clever tricks allows us to map the transverse field Ising Hamiltonian in $d$ dimensions to a classical anisotropic Ising Hamiltonian in $d+1$ dimensions \cite{Chakrabarti}. Let us demonstrate this in the case of the $1-D$ Quantum Ising model for simplicity. Its Hamiltonian simply reads

\begin{equation}
	H = -J\sum_{i=1}^N S_i^zS_{i+1}^z - \Gamma\sum_{i=1}^N S_i^x.
\end{equation}

The thermodynamical properties of this system can all be derived through its partition function $Z = \mathrm{Tr}\exp(-\beta H)$, with $\beta=1/k_BT$. Everything starts with the Trotter formula \cite{trotter}:

\begin{equation}
	\exp\left(A_1+A_2\right) =  \lim_{M\to\infty}\left[\exp\left(A_1/M\right)\exp\left(A_2/M\right)\right]^M.
\end{equation}

Writing our Hamiltonian as $H = H_0 + V$, where $H_0$ is the spin-spin interaction and $V$ the action of the magnetic field, one can expand the partition function as follows (keeping the large $M$ limit implicit for neatness):

\begin{align*}
	Z &= \mathrm{Tr}e^{-\beta\left(H_0+V\right)}\\
	&= \mathrm{Tr}\left[e^{-\beta H_0/M}e^{-\beta V/M}\right]^M\\
	&= \sum_{\left\{S^1_1,\cdots,S^1_N\right\}}\bra{S^1_1,\cdots,S^1_N}e^{-\beta H_0/M}e^{-\beta V/M}\cdots\\
	&\hspace{3cm}\times e^{-\beta H_0/M}e^{-\beta V/M}\ket{S^1_1,\cdots,S^1_N},
\end{align*}

where $\ket{S_i^1} = \ket{\pm1}_i$ is an eigenstates of $S_i^z$. The sum runs over all possible configurations for the lattice. Now, between each pair of exponential we can insert the identity in the form

\begin{equation}
	\mathds{1} = \sum_{\left\{S^k_1,\cdots,S^k_N\right\}}\ket{S^k_1,\cdots,S^k_N}\bra{S^k_1,\cdots,S^k_N}.
\end{equation}

The $k$ index, which simply labels one complete set of eigenstates of $S_z$, can be regarded as an additional dimension to our lattice (often referred to as the Trotter dimension). The spin $S_i^k$ can be understood the spin living on the $(i,k)$ site of a 2 dimensional lattice. Then

\begin{equation}
	Z = \sum_{\left\{S\right\}}\prod_{k=1}^M\bra{S^k_1,\cdots,S^k_N}e^{-\beta H_0/M}e^{-\beta V/M}\ket{S^{k+1}_1,\cdots,S^{k+1}_N},
\end{equation}

where we introduced periodic boundary conditions in the Trotter dimension, $\ket{S^{M+1}_1,\cdots,S^{M+1}_N} = \ket{S^1_1,\cdots,S^1_N}$ so as to match the braket sandwiching of the trace. We have use the shorthand $\sum_{\left\{S\right\}} = \sum_{\left\{S_1^1,\cdots,S_N^1\right\}}\cdots\sum_{\left\{S_1^M,\cdots,S_N^M\right\}}$, which is just a sum over all possible configurations of our 2D lattice. Now, the $H_0$ exponential is diagonal in the $\ket{S^k_1,\cdots,S^k_N}$ basis, so that, using its hermiticity

\begin{equation}
	\bra{S^k_1,\cdots,S^k_N}e^{-\beta H_0/M} = \exp\left[\frac{\beta J}{M}\sum_{i=1}^N S_i^k S_{i+1}^k\right]\bra{S^k_1,\cdots,S^k_N}.
\end{equation}

The second exponential now reads

\begin{align*}
	\bra{S^k_1,\cdots,S^k_N}e^{-\beta V/M}\ket{S^{k+1}_1,\cdots,S^{k+1}_N} &= \bra{S^k_1,\cdots,S^k_N}\exp\left[\frac{\beta\Gamma}{M}\sum_{i=1}^N S_i^x\right]\ket{S^{k+1}_1,\cdots,S^{k+1}_N}\\
	&= \prod_{i=1}^N\bra{S_i^k}e^{\beta\Gamma S_i^x/M}\ket{S_i^{k+1}}.
\end{align*}

Here comes a trick, keeping in mind that $S^x$ just flips the spin and that $(S^x)^2=\mathds{1}$:

\begin{align*}
	\bra{S}e^{aS^x}\ket{S'} &= \bra{S}\left(\sum_{n=0}^\infty\frac{a^{2n}(S^x)^{2n}}{n!} + \sum_{n=0}^\infty\frac{a^{2n+1}(S^x)^{2n+1}}{n!}\right)\ket{S'}\\
	&= \cosh{a}\braket{S,S'} + \sinh{a}\bra{S}S^x\ket{S'}\\
	&= \cosh{a}\braket{S,S'} + \sinh{a}\braket{S,-S'}\\
	&= \begin{cases}
		&\cosh{a} \text{ if } S=S',\\
		&\sinh{a} \text{ if } S=-S'.
	\end{cases}
\end{align*}

Now, simply note that

\begin{align*}
	\left(\frac{1}{2}\sinh{2a}\right)^{1/2}\exp\left[\frac{SS'}{2}\ln\coth{a}\right] &= \sqrt{\sinh{a}\cosh{a}}\times
		\begin{cases}
			&\exp\left[\dfrac{1}{2}\ln\coth{a}\right] \text{ if } S=S'			\vspace{1mm}\\
			&\exp\left[-\dfrac{1}{2}\ln\coth{a}\right] \text{ if } S=-S'
		\end{cases}\\
	& = \sqrt{\sinh{a}\cosh{a}}\times
	\begin{cases}
		&\sqrt{\cosh{a}/\sinh{a}} \text{ if } S=S'\\
		&\sqrt{\sinh{a}/\cosh{a}} \text{ if } S=-S'
	\end{cases}\\
	&= \begin{cases}
		&\cosh{a} \text{ if } S=S'\\
		&\sinh{a} \text{ if } S=-S'
	\end{cases}\\
	&= \bra{S}e^{aS^x}\ket{S'}
\end{align*}

This was a long road, but we can finally put everything together.

\begin{align*}
	Z &= \sum_{\left\{S\right\}}\prod_{k=1}^M\prod_{i=1}^N\exp\left[\frac{\beta J}{M}S_i^kS_{i+1}^k\right]\left(\frac{1}{2}\sinh{2\beta\Gamma/M}\right)^{1/2}\exp\left[\frac{S_i^kS_i^{k+1}}{2}\ln\coth(\beta\Gamma/M)\right]\\
	&= \left(\frac{1}{2}\sinh(2\beta\Gamma/M)\right)^{MN/2}\sum_{\left\{S\right\}}\exp\left[-\frac{\beta}{M}\left(-J\sum_{i=1}^N\sum_{k=1}^M S_i^k S_{i+1}^k - \frac{M}{2\beta}\ln\coth\left(\dfrac{\beta\Gamma}{M}\right)\sum_{i=1}^N\sum_{k=1}^M S_i^k S_i^{k+1}\right)\right]\\
	&= C\hspace{1mm}\mathrm{Tr}\left[e^{-\beta_\mathrm{cl}H_\mathrm{eff}}\right],
\end{align*}

with the classical temperature $\beta_\mathrm{cl} = \beta/M$. Let us also abbreviate $K_M = \dfrac{1}{2\beta}\ln\coth\left(\dfrac{\beta\Gamma}{M}\right)$. In the large $M$ limit, the prefactor vanishes, so that it drops out of any physical observables (obtained by differentiating the logarithm of the partition function). One recognizes in $H_\mathrm{eff}$ the Hamiltonian of a classical (since the $S_i^k$ are just numbers) anisotropic Ising model without any magnetic field. The weird thing is that the coupling constant of the spin-spin interactions depend on the lattice size in the Trotter dimension and even on the temperature for the interactions along the Trotter dimension. The quantum Ising model dual to our $\mathds{Z}_2$ gauge Hamiltonian is two dimensional, the above result therefore needs to be generalized to higher dimensions. The quantum Ising model on a two dimensional $N_x\times N_y$ lattice

\begin{equation}
	H = -J\sum_{i,j}S_{i,j}\left(S^z_{i+1,j} + S^z_{i,j+1}\right) -\Gamma\sum_{i,j}S^x_{i,j},
\end{equation}

can be mapped to the 3D classical anisotropic Ising model on a $N_x\times N_y \times N_z$ lattice

\begin{equation}
	H_\mathrm{eff} = -\sum_{k=1}^{N_z}\sum_{i,j}\left[J S_{i,j,k}\left(S_{i+1,j,k} + S_{i,j+1,k}\right) + K_{N_z} S_{i,j,k}S_{i,j,k+1}\right]
\end{equation}


\section{Metropolis-Hasting Monte-Carlo algorithm}

We have derived a classical Ising Hamiltonian which shares, up to a constant, the same partition function with the quantum model. We should therefore be able to study its thermodynamical properties, mainly its phase transitions, through a simple Monte-Carlo algorithm, which naively works as follows:

\begin{tcolorbox}[title=Metropolis-Hasting algorithm]
	
	\begin{enumerate}
		
		\item Generate a random configuration.
		
		\item Spin one randomly chosen spin on the lattice and compute the change in energy $dE$.
			\begin{itemize}
				\item If $dE<0$: accept this new configuration.
				
				\item Else: accept the configuration with probability $\exp\left(-\beta dE\right)$, otherwise reject it.
			\end{itemize}
		
		\item Repeat until the required number of configurations have been generated.
		
		\item Ditch the first $N_\mathrm{eq}$ generated configurations before reaching equilibrium
		
	\end{enumerate}
	
\end{tcolorbox}

The Metropolis procedure was specifically designed to generate a sample configurations according to their probability distribution. One can then compute averages without worrying about computing the partition function of the system:

\begin{equation}
	\langle O\rangle = \frac{1}{N_\mathrm{sample}}\sum_\mathrm{sample} O.
\end{equation}

A few subtleties have to be taken into account. The time taken by the Metropolis algorithm to reach equilibrium -that is to say the number of steps required to generate configurations with the proper probability distribution- is not easy to guess, and has to be adjusted "by hand" quite qualitatively. One can for example plot the evolution of desired observables as the Monte-Carlo sampling goes on. We can however expect typical behaviours.

\begin{itemize}
	
	\item The bigger the lattice, the more steps will be required to thoroughly explore the space of configurations (of which there are $2^{N_xN_yN_z}$!). This is troublesome in our case since the classical mapping only holds in the large $N_z$ limit. This can be partially solved by the means of cluster methods, where the Metropolis update consists of flipping whole chunks of spins at the same time. Computing the change in energy will however naively requires going through the whole lattice, with complexity $\mathcal{O}\left(N_xN_yN_z\right)$, which might be too costly for thousands of Metropolis updates.
	
	\item Disordered phases (for example at high temperatures) will naively require less steps to reach equilibrium, since we can expect the initial random configuration to quite likely with respect to the actual distribution.
	
\end{itemize}


\section{Results}

\subsection{Phase transitions}

\section{Closing remarks and perspectives}


\clearpage\printbibliography

\end{document}